---
title: "Fire & Flood Risks in EJ Communities"
author: "David Amanfu, Chayan Jani, Kelsie Roberton"
date: "April 2022"
output: 
  html_document:
    toc: true
    toc_float: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r, initialize}
# Set your working directory
getwd()
#knitr::opts_knit$set(root.dir = '/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Project/')
knitr::opts_chunk$set(echo = TRUE)
```

<!-- ```{=tex} -->

<!-- \newpage -->

<!-- \tableofcontents  -->

<!-- \newpage -->

<!-- \listoftables  -->

<!-- \newpage -->

<!-- \listoffigures  -->

<!-- \newpage -->

<!-- ``` -->

##Initializing Setup

```{r setup, include=FALSE}
# Load your packages
library(agricolae)
library(corrplot)
library(cowplot)
library(dataRetrieval)
library(extrafont)
library(extrafontdb)
library(ggpubr)
library(ggthemes)
library(hrbrthemes)
library(Kendall)
library(leaflet)
library(lubridate)
library(mapview)
library(rvest)
library(sf)
library(tidycensus)
library(tidyverse)
library(trend)
library(tseries)
library(viridis)
library(zoo)


# Set your ggplot theme
#Disable on-the-fly projections
sf::sf_use_s2(FALSE)
#Fix Mapview
mapviewOptions(fgb = FALSE)

Theme2 <- theme_ipsum()+
  theme(legend.position = "bottomYes",
        legend.key = element_rect(fill = "white", colour = "black"),legend.direction = "horizontal",
        legend.title = element_text(face = "bold"))
theme_set(Theme2)
```

#Data Setup
##Data Retrieval

For our analysis, we used data collected from the Environmental
Protection Agency (EPA) EJ Screen to look at pre-1960 housing and
low-income housing populations. For reference, EJScreen is an
environmental justice screening and mapping tool that provides the EPA
with datasets on a national-scale that combines environmental and
geographic indicators (epa.gov). All EJScreen indicators are publicly
available data on the EPA website. For our analysis, data was downloaded
as an Excel file [here](https://gaftp.epa.gov/EJSCREEN/). Our analysis
used 2020 data from EJScreen as opposed to 2021 data in order to avoid
any data inconsistencies that may have not been resolved in the 2021
data. As well, the Census Bureau's data we utilized was from 2020, and
for consistency sake we continued with 2020 data. We are not aware of
any significant changes between the two otherwise. To nicely overlay the
information from the census tracts onto a geometry, we included matching
census tract information from the US Census Bureau's Gazetteer Files
database. The Gazetteer files, downloadable
[here](https://www.census.gov/geographies/reference-files/time-series/geo/gazetteer-files.2020.html),
include geographic identifying information including tract tags, county
names, and importantly, representative latitude and longitude
coordinates. To gather data for wildfire and flood risk, we utilized
FEMA's National Risk Index (NRI) data. The NRI data contains information
regarding a geographic boundary's exposure to and risk implications of
18 natural hazards. The data is made available at both the county and
census tract level. We used the csv for all census tracts, downloadable
[here](https://hazards.fema.gov/nri/data-resources), due to the large
size of the shapefile. We combined this data with US Census shapefiles
at the census-tract level, with Cartographic Boundary Files,
downloadable \[here\]. ## Load Datasets

```{r load datasources, include=FALSE}
# Load your datasets
home_path <- '~/Desktop/Duke MPP/Environ Data /872 Final Project'

NRINorCar <- read.csv(paste(home_path,'/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/NRI_Table_CensusTracts_NorthCarolina/NRI_Table_CensusTracts_NorthCarolina.csv',sep=""), stringsAsFactors = TRUE)
NRIFlor <- read.csv(paste(home_path,'/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/NRI_Table_CensusTracts_Florida/NRI_Table_CensusTracts_Florida.csv',sep=""), stringsAsFactors = TRUE)
EJScreen20 <- read.csv(paste(home_path,'/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/EJSCREEN_2020_StatePctile.csv',sep=""),stringsAsFactors = TRUE)
```

#Data Wrangling

We started our analysis by selecting the following columns from the raw
(name of data file) GEOID, NAMESLAD, NAMESLADCO, ALAND, AWATER, geometry
for each state (NC, FL and CA). Next, we filtered for each state that
has to do with fire and floods. Following this, EJScreen data came in
census block groups. We used the 'group by' summarize function to
aggregate into census tracts. Following this, we can match these census
tracts by the NRI risk data files (by state).

Because EJ Screen data does not come with geometric coordinates, we had
to import and merge the census gazetteer files to at least give them a
corresponding point within the census. Following this, we attempted to
merge both of the filtered files for each state's fire and flood data.
Finally, we were able to create three data-sets per shape: fire-risks,
flood-risks, and EJScreen demographic. We then proceeded with the
exploratory analysis, which can be shown below

Given the extensive nature of all of the datasets, we decided to remove
California from our analyses and focus on North Carolina and Florida.
##Filter Steps 
### Initial Filter By State

```{r initial filter}
NRIFlor2 <- NRIFlor %>%
  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,
         ends_with("RATNG"),ends_with("RISKS"),ends_with("RISKR"),ends_with("EVENTS"),
         ends_with("SCORE"),ends_with("EALS"),ends_with("EALR")) %>%
         select(!starts_with("LTNG"),!starts_with("AVLN"),!starts_with("HAIL"),
                !starts_with("LNDS"),!starts_with("VLCN"))

NRINorCar2 <- NRINorCar %>%
  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,
         ends_with("RATNG"),ends_with("RISKS"),ends_with("RISKR"),ends_with("EVENTS"),
         ends_with("SCORE"),ends_with("EALS"),ends_with("EALR")) %>%
         select(!starts_with("LTNG"),!starts_with("AVLN"),!starts_with("HAIL"),
                !starts_with("LNDS"),!starts_with("VLCN"))

EJScreen20 <- EJScreen20 %>%
  filter(grepl('Florida|North Carolina',STATE_NAME))
```

###Filter NRI Dataset For Risks of Interest

```{r NRI risk filters}
#flood risks
NRIFlorFlood <- NRIFlor2 %>%
  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,starts_with("CFLD"),starts_with("RFLD"))
NRINorCarFlood <- NRINorCar2 %>%
  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,starts_with("CFLD"),starts_with("RFLD"))

#fire risks
NRIFlorFire <- NRIFlor2 %>%
  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,starts_with("WFIR"))
NRINorCarFire <- NRINorCar2 %>%
  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,starts_with("WFIR"))
```

```{r EJScreen Filters}
EJ_filter <- EJScreen20 %>%
  select(ID,STATE_NAME,contains("INCPCT"),contains("HSPCT"),LINGISOPCT,PRE1960PCT,contains("LDPNT"),contains("PM25"))
```

```{r EJScreen State Filters}
EJ_NorCar <- EJ_filter %>%
  filter(STATE_NAME=='North Carolina')

EJ_Florida <-EJ_filter %>%
  filter(STATE_NAME=="Florida")
```

```{r Optional Read/Write EJScreen Files,echo=FALSE, eval=FALSE}
## COMMENT THESE OUT THEN READ THEM IN
# write.table(EJ_NorCar,'./Potential Data Sources/Census Tracts/EJScreen_NorthCarolina.csv',col.names=TRUE,sep=",")
# write.table(EJ_Florida,'./Potential Data Sources/Census Tracts/EJScreen_Florida.csv',col.names=TRUE,sep=",")
# ###

# EJ_NorCar <- EJScreen20 %>%
#   filter(STATE_NAME=='North Carolina')
# 
# EJ_Florida <-EJScreen20 %>%
#   filter(STATE_NAME=="Florida")

#write.csv(EJ_NorCar,'./Potential Data Sources/Census Tracts/EJScreen_NC.csv',col.names)
#write.csv(EJ_Florida,'./Potential Data Sources/Census Tracts/EJScreen_FLorida.csv')
```

##Adding Location Data

###Shape Files from Census Data

Add in Shape Files
<https://www.census.gov/geographies/mapping-files/time-series/geo/cartographic-boundary.html>

```{r census tracts}
california_tract <- st_read('./Potential Data Sources/Census Tracts/cb_2020_cali_tract_500k/cb_2020_06_tract_500k.shp',)
norcar_tract <- st_read('./Potential Data Sources/Census Tracts/cb_2020_norcar_tract_500k/cb_2020_37_tract_500k.shp')
florida_tract <- st_read('./Potential Data Sources/Census Tracts/cb_2020_flor_tract_500k/cb_2020_12_tract_500k.shp')
#mapview(california_tract,zcol="AWATER")
```

Using Point Data for Longitude/Latitude Coordinates from US Census
Gazetteer Files
<https://www.census.gov/geographies/reference-files/time-series/geo/gazetteer-files.2020.html>

```{r Census Tract Point Long/Lat Coordinates}
#https://www.census.gov/geographies/reference-files/time-series/geo/gazetteer-files.2020.html
gazetteer <- read.delim('./Potential Data Sources/Census Tracts/2021_Gaz_tracts_national.txt',stringsAsFactors = TRUE)

fl_gazetteer <- gazetteer %>%
  filter(USPS=="FL")

fl_gazetteer <- fl_gazetteer %>%
  select(!c(ALAND_SQMI, AWATER_SQMI)) %>%
  mutate(TRACTID = fl_gazetteer$GEOID)

nc_gazetteer <- gazetteer %>%
  filter(USPS=="NC")

nc_gazetteer <- nc_gazetteer %>%
  select(!c(ALAND_SQMI, AWATER_SQMI)) %>%
  mutate(TRACTID = nc_gazetteer$GEOID)
```

###Filtering With Shape Geometry

```{r florida file filtering}
florida_tract <- florida_tract %>%
  select(GEOID,NAMELSAD, NAMELSADCO, ALAND, AWATER, geometry) %>%
  mutate(TRACTFIPS = as.double(substr(florida_tract$GEOID, 1,15)))
NRIFlorFire <-NRIFlorFire %>%
  select(TRACTFIPS,POPULATION,starts_with ("WFIR"))
NRIFlorFlood <- NRIFlorFlood %>%
  select(TRACTFIPS,POPULATION,starts_with("CFLD"),starts_with("RFLD"))


# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_Florida.csv'
# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_Florida.csv'
# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_Florida.csv'
EJScreen_Florida <- EJ_Florida
names(EJScreen_Florida)[names(EJScreen_Florida) == 'ID'] <- 'BLOCKID' 
EJScreen_Florida <- EJScreen_Florida %>% mutate(TRACTID = as.double(substr(EJScreen_Florida$BLOCKID,1,11)))

         
EJ_Flor <- EJScreen_Florida %>%
select(!starts_with("T_")) %>%
group_by(TRACTID) %>%
summarize(across(!where(is.factor), mean))
```

```{r north carolina file filtering}
norcar_tract <- norcar_tract %>%
  select(GEOID,NAMELSAD, NAMELSADCO, ALAND, AWATER, geometry) %>%
  mutate(TRACTFIPS = as.double(substr(norcar_tract$GEOID, 1,15)))
NRINorCarFire <-NRINorCarFire %>%
  select(TRACTFIPS,POPULATION,starts_with ("WFIR"))
NRINorCarFlood <- NRINorCarFlood %>%
  select(TRACTFIPS,POPULATION,starts_with("CFLD"),starts_with("RFLD"))


# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_NorthCarolina.csv'
# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_NorthCarolina.csv'
# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_NorthCarolina.csv'
EJScreen_norcar <- EJ_NorCar
names(EJScreen_norcar)[names(EJScreen_norcar) == 'ID'] <- 'BLOCKID' 
EJScreen_norcar <- EJScreen_norcar %>% mutate(TRACTID = as.double(substr(EJScreen_norcar$BLOCKID,1,11)))

EJ_NorCar <- EJScreen_norcar %>%
select(!starts_with("T_")) %>%
group_by(TRACTID) %>%
summarize(across(!where(is.factor), mean))
```

###Merging NRI Data with Tract Shapes, EJScreen with Tract Coordinates

```{r attempts to merge}
#Florida
FL_NRI_fire_shp <- left_join(florida_tract,NRIFlorFire,by="TRACTFIPS")
FL_NRI_flood_shp <- left_join(florida_tract,NRIFlorFlood,by="TRACTFIPS")

FL_EJ_df <- left_join(fl_gazetteer,EJ_Flor,by="TRACTID")
FL_EJ_shp <- FL_EJ_df %>%
  st_as_sf(coords = c('INTPTLONG','INTPTLAT'), crs=4269)


#North Carolina
NC_NRI_fire_shp <- left_join(norcar_tract,NRINorCarFire,by="TRACTFIPS")
NC_NRI_flood_shp <- left_join(norcar_tract,NRINorCarFlood,by="TRACTFIPS")

NC_EJ_df <- left_join(nc_gazetteer,EJ_NorCar,by="TRACTID")
NC_EJ_shp <- NC_EJ_df %>%
  st_as_sf(coords = c('INTPTLONG','INTPTLAT'), crs=4269)
```

#Data Exploration 
##Simple Mapping
```{r simple mapview tests}
#FL
mapview(FL_NRI_flood_shp,zcol="RFLD_RISKR")
mapview(FL_EJ_shp,zcol="LOWINCPCT")
# mapview(FL_NRI_fire_shp,zcol="POPULATION")
mapview(FL_NRI_fire_shp,zcol="POPULATION")+ mapview(FL_NRI_flood_shp,zcol="RFLD_RISKR") +mapview(FL_EJ_shp,zcol="LOWINCPCT")

#NC
mapview(NC_NRI_flood_shp,zcol="RFLD_RISKR")
mapview(NC_EJ_shp,zcol="LOWINCPCT")
# mapview(NC_NRI_fire_shp,zcol="POPULATION")
mapview(NC_NRI_fire_shp,zcol="POPULATION")+mapview(NC_NRI_flood_shp,zcol="RFLD_RISKR")+mapview(NC_EJ_shp,zcol="LOWINCPCT")
```

## Filtering Out Missing Data
```{r Dropping Incomplete Info}
FL_EJ_shp <- FL_EJ_shp %>%
  drop_na(BLOCKID)
NC_EJ_shp <- NC_EJ_shp %>%
  drop_na(BLOCKID)

FL_NRI_fire_shp <- FL_NRI_fire_shp %>%
  drop_na(WFIR_RISKS)   #%>%filter(WFIR_RISKS>0)
NC_NRI_fire_shp <- NC_NRI_fire_shp %>%
  drop_na(WFIR_RISKS)   #%>%filter(WFIR_RISKS>0)

FL_NRI_flood_shp <- FL_NRI_flood_shp %>%
  drop_na(POPULATION)
NC_NRI_flood_shp <- NC_NRI_flood_shp %>%
  drop_na(POPULATION)
```

#Data Analysis

## Simple Correlative Graphs

```{r simple ggplots FL}
ggplot(FL_EJ_df,aes(y=LOWINCPCT,x=PM25))+geom_point(aes(color=LINGISOPCT))+geom_smooth(method="lm")+labs(title="PM25 v Low Income %",y="% Low Income Census Tract",x="PM25 Concentration")

ggplot(FL_NRI_fire_shp,aes(y=POPULATION,x=WFIR_RISKR))+geom_boxplot()+coord_flip()+labs(title="Fire Risk versus Population",y="Fire Risk")
ggplot(FL_NRI_flood_shp,aes(y=POPULATION,x=CFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Coastal Flood Risk versus Population",y="Coastal Flood Risk")
ggplot(FL_NRI_flood_shp,aes(y=POPULATION,x=RFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Riverine Flood Risk versus Population",y="Riverine Flood Risk")
```

```{r simple ggplots NC}
ggplot(NC_EJ_df,aes(y=LOWINCPCT,x=PM25))+geom_point(aes(color=LINGISOPCT))+geom_smooth(method="lm")+labs(title="PM25 v Low Income %",y="% Low Income Census Tract",x="PM25 Concentration")

ggplot(NC_NRI_fire_shp,aes(y=POPULATION,x=WFIR_RISKR))+geom_boxplot()+coord_flip()+labs(title="Fire Risk versus Population",y="Fire Risk")
ggplot(NC_NRI_flood_shp,aes(y=POPULATION,x=CFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Coastal Flood Risk versus Population",y="Coastal Flood Risk")
ggplot(NC_NRI_flood_shp,aes(y=POPULATION,x=RFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Riverine Flood Risk versus Population",y="Riverine Flood Risk")
```

##Visualized Mapped Graphics

```{r mapview tests FL}
mapview(list(FL_NRI_fire_shp,FL_EJ_shp),
        zcol=c("WFIR_RISKS","LOWINCPCT"),
        layer.name=list("Fire Risk","Share of Low Income Households"))

mapview(list(FL_NRI_flood_shp,FL_EJ_shp),
        zcol=list("CFLD_RISKS","LOWINCPCT"),
        layer.name=list("Coastal Flood Risk","Share of Low Income Households"),
        col.regions=list(mapviewPalette("mapviewVectorColors"),mapviewPalette("mapviewTopoColors")),
        alpha.regions=.5)

# mapview(FL_NRI_fire_shp,zcol="WFIR_RISKS")
# mapview(FL_EJ_shp,zcol="LOWINCPCT",alpha.regions=.5,col.regions= mapviewPalette("mapviewTopoColors"))
#mapview(FL_EJ_shp,zcol="PRE1960PCT")

# mapshot(m, file = paste0(getwd(), "/map.png"),
#           remove_controls = c("homeButton", "layersControl"))
```

```{r mapview tests NC}
mapview(list(NC_NRI_fire_shp,NC_EJ_shp),
        zcol=c("WFIR_RISKS","LOWINCPCT"),
        layer.name=list("Fire Risk","Share of Low Income Households"))

mapview(list(NC_NRI_flood_shp,NC_EJ_shp),
        zcol=list("CFLD_RISKS","LOWINCPCT"),
        layer.name=list("Coastal Flood Risk","Share of Low Income Households"),
        col.regions=list(mapviewPalette("mapviewVectorColors"),mapviewPalette("mapviewTopoColors")),
        alpha.regions=.5)

# mapview(NC_NRI_fire_shp,zcol="WFIR_RISKS")
# mapview(NC_EJ_shp,zcol="LOWINCPCT",alpha.regions=.5,col.regions= mapviewPalette("mapviewTopoColors"))
#mapview(NC_EJ_shp,zcol="PRE1960PCT")

# mapshot(m, file = paste0(getwd(), "/map.png"),
#           remove_controls = c("homeButton", "layersControl"))
```

##Statistical Analyses 
### Merged Datasets

```{r}
#new dataset for low income population from above tractfips - FL, NC
Lowincomepop <-rbind(EJ_Flor, EJ_NorCar)
Lowincomepop <-Lowincomepop%>%
  select(TRACTID:LOWINCPCT, P_LWINCPCT)
names(Lowincomepop)[names(Lowincomepop) == 'TRACTID'] <- 'TRACTFIPS' 

#final data set for fire risk
Firerisk <- rbind(NRIFlorFire, NRINorCarFire)
Firerisk<- Firerisk%>%
  filter(WFIR_RISKR == "Relatively High"| WFIR_RISKS == "Very High")
 
#final data set for flood risk - all 3 states with relatively high and high risk
Floodrisk <-rbind(NRIFlorFlood, NRINorCarFlood)
FLoodrisk<-Floodrisk%>%
  select(TRACTFIPS:CFLD_RISKR,RFLD_RISKS,RFLD_RISKS)
 
CFLoodrisk <- Floodrisk%>%
  filter(CFLD_RISKR == "Relatively High"| CFLD_RISKR == "Very High")
 
RFLoodrisk <- Floodrisk%>%
  filter(RFLD_RISKR == "Very High"| RFLD_RISKR == "Relatively High" )
```

### Coastal Flooding Tests

The p-value of the test is 0.01496, which is less than the significance
level alpha = 0.05. We can conclude that low income population and
coastal flood risks are significantly correlated with a correlation
coefficient of 0.1806238 and p-value of 0.01496.It has a positive
correlation which means flood risk is impacting a significant %of low
income population.

```{r}
#statistical test for coastal flooding and low income population
LI_CF_risk <- left_join(CFLoodrisk,Lowincomepop,by="TRACTFIPS")
 
LI_CF_risk.regression <-lm (LI_CF_risk$CFLD_RISKS~ LI_CF_risk$LOWINCPCT)
summary(LI_CF_risk.regression)
#comments
cor.test(LI_CF_risk$CFLD_RISKS,LI_CF_risk$LOWINCPCT)
#comments
 
par(mfrow = c(2,2), mar=c(4,4,4,4))
plot(LI_CF_risk.regression)
par(mfrow = c(1,1))
 
#mapview for coastal flooding - ensure population is represented by size of the dot
#comments
```

###Riverine Flooding Tests The p-value of the test is 0.4548, which is
greater than the significance level alpha = 0.05. From this We can
conclude that low income population and river flood risks are not
significantly correlated and correlation coefficient is -0.02787792,
negative implying inverse relationship.This might be due to the states
selected and the overall chances of river flooding. This also means that
the population at risk from river floods is not mostly low income.

```{r}
#statistical test for river flooding and low income population
LI_RF_risk <- left_join(RFLoodrisk,Lowincomepop,by="TRACTFIPS")
 
LI_RF_risk.regression <-lm (LI_RF_risk$RFLD_RISKS~ LI_RF_risk$LOWINCPCT)
summary(LI_RF_risk.regression)
#comments
cor.test(LI_RF_risk$RFLD_RISKS,LI_RF_risk$LOWINCPCT)
#comments
 
par(mfrow = c(2,2), mar=c(4,4,4,4))
plot(LI_RF_risk.regression)
par(mfrow = c(1,1))
 
#mapview for river flooding - ensure population is represented by size of the dot
#comments
```

###Wildfire Risk Tests 
The p-value of the test is 0.53, which is greater than the significance level alpha = 0.05. From this we can conclude that low income population and wildfire risks are not significantly correlated and correlation coefficient is -0.03065284 , negative implying inverse relationship.This might be due to the states selected and the overall chances of wildfire risk. This also means that the population at risk from wildfire is not mostly low income.

```{r}
#statistical test for wildfire and low income population
LI_WF_risk <- left_join(Firerisk,Lowincomepop,by="TRACTFIPS")
 
LI_WF_risk.regression <-lm (LI_WF_risk$WFIR_RISKS~ LI_WF_risk$LOWINCPCT)
summary(LI_WF_risk.regression)
#comments
cor.test(LI_WF_risk$WFIR_RISKS,LI_WF_risk$LOWINCPCT)
#comments
 
par(mfrow = c(2,2), mar=c(4,4,4,4))
plot(LI_WF_risk.regression)
par(mfrow = c(1,1))
 
#mapview for wildfire - ensure population is represented by size of the dot
#comments
 
```

#Conclusions 
The project aimed to look at the Environment Justice component from natural hazards. While there were several EJ communities, for our final report we focused on the % of the low income population. We looked at the National risk index for three natural hazard categories: coastal flooding, river flooding and wildfire. We looked at these parameters for two states: Florida and North Carolina

On analyzing the counties with high or relatively high natural hazard risk and comparing it with the % of low income population in that county we found that: 
- In all three states where there was a high risk of coastal flooding, there was a significant percentage of low income population. 
- In all three states where there was high risk of river flooding and wildfire risk there was not a significant percentage of low income population.

##Limitations 
The datasets we decided to use were very dense, making it difficult for R to adequately download the values. Unfortunately, a lot of filtering for the original EJScreen dataset was required to actually upload, push, and pull the data from GitHub to R and vice versa. This left a lot of precious time that could have been spent exploring air quality, which was another natural hazard risk that we initially planned to analyze.

Furthermore, there was a lot of data for the state of California compared to North Carolina and Florida. That said, it made it difficult for us to understand and identify a trend. We removed California from the study as a result.

The data also had many categorical variables for NRI data and different populations for EJ data. This leaves a lot of scope for different analysis and finding statistically significant results.

